---
title: "Lasso Regression"
author: "Cem Kalender (2920734)"
date: "10/13/2021"
output: html_document
---

```{r}
library(ISLR)
library(MASS)
library(tidyverse)
library(glmnet)

df_train <- read_rds("train.rds")
df_test  <- read_rds("test.rds")
```

We will perform a Laso regression due to a high number of p L1 regularization allows for shrinking unassociated parameters to zero.

```{r}
#Before proceeding ensure that the missing values have been removed from the data----

any(is.na(df_train)) # No NaNs
```

```{r}
#making matrix of our dataset
x <- df_train[,-31]
x <- model.matrix(score ~., df_train)[,]
dim(x)

y <- df_train$score

```

splitting training and validation set:
```{r}

set.seed(123)
train <- sample(1:nrow (x), size = nrow(x)*0.75)
valid <- (-train)
y.valid <- y[valid]

```

# defining lasso model and training it

``` {r}

lasso.mod <- glmnet(x[train , ], y[train], alpha = 1) #default lambda values
dim (coef(lasso.mod))

plot(lasso.mod) #coefficients against L1 regularization

```


# We now perform cross-validation and compute the associated test error.
```{r}
set.seed (1)
cv.out <- cv.glmnet(x[train, ], y[train], alpha = 1)

plot (cv.out) #plotting the lambda values against MSE

#λ chosen by cross-validation

bestlam <- cv.out$lambda.min #extracting lambda value for lowest MSE by using CV

#predicting on the validation set:
lasso.pred <- predict(lasso.mod , s = bestlam ,
                         newx = x[valid , ])



#these are coefficients by the way:
out <- glmnet (x[train, ], y[train], alpha = 1,
                     lambda = bestlam)

coefs <- predict(out, type = "coefficients",
         s = bestlam)[1:31, ]

which(coefs != 0)
```

# lets calculate MSE on validation set

```{r}
#calculating rMSE:
mean((lasso.pred - y.valid)^2) #rMSE is 0.7342965

#vector of predicted values for score:
lasso.pred

```

# now lets estimate the scores from the test dataset

```{r}

#first lets use model.matrix to convert our categorical variables to dummy variables.

x_test <- model.matrix(~.,data=df_test)[,]
dim(x_test)

#predicting on test data:
lasso.predtest <- predict(lasso.mod , s = bestlam ,
                       newx = x_test)


lasso.predtest #predictions
```

Plotting the distribution of the predicted test values:
```{r}
hist(lasso.predtest, breaks=15) #distribution of predictions

```